---
title: 'COVID-19 ANALYSIS'
author: "Palka, Matthew"
output:
html_document: default
pdf_document: default
---
```{r setup, include = FALSE}
# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)

```
Project Goals:
  
  1)Descriptive Statistics: Do an Exploratory Data Analysis (EDA) and provide 
appropriate summary statistics / visualizations to help understand the spread
of the disease (incidence) as well as its fatality rate. 

2)Inferential Statistics: Pick 2 countries and compare their incidence and
fatality rates using hypothesis testing and confidence interval methods

3)Correlation: Pick all the countries and evaluate the relationship between 
incidence rates and fatality rates. Compute the correlation coefficient, if relevant.

4)Regression: Pick United States. Look at the time series of cases and 
time series of deaths. Use a regression model to predict the number of cases 
and the number of deaths for the next 5 days in the future.

Final report: Submit a pdf file with your analysis results and conclusions.

```{r}
install.packages("ggplot2", repos = "http://cran.us.r-project.org")
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("funModeling", repos = "http://cran.us.r-project.org")
install.packages("Hmisc", repos = "http://cran.us.r-project.org")
install.packages("ggpubr", repos = "http://cran.us.r-project.org")
install.packages("dplyr", repos = "http://cran.us.r-project.org")
install.packages("maps", repos = "http://cran.us.r-project.org")
install.packages("GGally", repos = "http://cran.us.r-project.org")
library(GGally)
library(readxl)
library(funModeling) 
library(tidyverse) 
library(Hmisc)
library(ggplot2)
library(gridExtra)

```

```{r}
mydata <- read_excel("COVID-19-geographic-disbtribution-worldwide-2020-12-14.xlsx", 
     col_types = c("date", "skip", "numeric", 
         "skip", "numeric", "numeric", "text", 
         "text", "skip", "numeric", "text", 
         "numeric"))
print(mydata)
```

```{r}
describe(mydata)
ttlsbyCountry <- mydata %>%
group_by(countriesAndTerritories, continentExp, month) %>%
summarise(deaths = sum(deaths),cases = sum(cases), mortality = (deaths/cases))

ttlsbyContin <- mydata %>%
group_by(continentExp, month) %>%
summarise(deaths = sum(deaths), cases = sum(cases), mortality = (deaths/cases))

newcolordeaths <- qplot(dateRep, deaths, colour = continentExp, 
   data = mydata)
newcolorcases <- qplot(dateRep, cases, colour = continentExp, 
   data = mydata)
newcolordeaths
newcolorcases

test1 <- qplot(cases, deaths, colour = continentExp, 
   data = ttlsbyCountry)+
   facet_wrap( ~ month )
test2 <- qplot(month, cases, colour = continentExp, 
   data = ttlsbyContin)+
   geom_line()
test1
test2

gDeath <- ggplot(mydata, aes(x=dateRep, y=deaths)) +
  geom_line() +
  xlab("month") + 
  facet_grid(cols = vars(continentExp))
gDeath
gptDeath <- ggplot(mydata, aes(dateRep,deaths,color = continentExp)) +
      geom_smooth(method=lm)
gptDeath

gCases <- ggplot(mydata, aes(x=dateRep, y=cases)) +
  geom_line() +
  xlab("date")

gBox <-  ggplot(mydata, aes(x=continentExp, y=deaths)) +
          geom_boxplot(outlier.colour="red", outlier.shape=16,
             outlier.size=2, notch=FALSE)


#grid.arrange(newcolordeaths,newcolorcases, ncol = 2, nrow = 1)




```
```{r}
#Part 2
#2)Inferential Statistics: Pick 2 countries and compare their incidence and
#fatality rates using hypothesis testing and confidence interval methods
#Compare Poland and Russia, Null Hypothesis that both countries were seeing a equal spike in cases as well as Deaths

polData <- mydata %>%
  filter(countriesAndTerritories=="Poland")%>%
  group_by(month,countriesAndTerritories)%>%
  summarise(deaths = sum(deaths),cases = sum(cases), mortality = (deaths/cases))
rusData <- mydata %>%
  filter(countriesAndTerritories=="Russia")%>%
  group_by(month,countriesAndTerritories)%>%
  summarise(deaths = sum(deaths),cases = sum(cases), mortality = (deaths/cases))
t.test(polData$mortality,rusData$mortality)

#Based off the results of the T-test wee will most likely reject the null hypothesis that both Poland and Russians mortality rates are the same since the P-value of .029 falls below the .05 threshold so we are able to reject the null hypothesis.

```
```{r}
#3)Correlation: Pick all the countries and evaluate the relationship between 
#incidence rates and fatality rates. Compute the correlation coefficient, if relevant.
ttlCases = mydata$cases
ttlDeaths = mydata$deaths
cor(ttlCases,ttlDeaths)

#Overall there is a strong correlation between having more cases and having more deaths.
```
```{r}
#4)Regression: Pick United States. Look at the time series of cases and 
#time series of deaths. Use a regression model to predict the number of cases 
#and the number of deaths for the next 5 days in the future.

usData <- mydata %>%
  filter(countriesAndTerritories=="United_States_of_America")%>%
  group_by(dateRep,countriesAndTerritories)%>%
  summarise(deaths = sum(deaths),cases = sum(cases), mortality = (deaths/cases))
usData$index <- 1:nrow(usData)
usData

fit_1 <- lm(deaths ~ index, data = usData)
summary(fit_1)
fit_2 <- lm(cases ~ index, data = usData)
summary(fit_2)

ggplot(data=usData, aes(fit_1$residuals)) +
geom_histogram(binwidth = 1, color = "black", fill = "purple4") +
theme(panel.background = element_rect(fill = "white"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Histogram for Model Residuals")



ggplot(data = usData, aes(x = dateRep, y = deaths)) +
geom_point() +
stat_smooth(method = "lm", col = "dodgerblue3") +
theme(panel.background = element_rect(fill = "white"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Linear Model Fitted to Death Data")

ggplot(data = usData, aes(x = dateRep, y = cases)) +
geom_point() +
stat_smooth(method = "lm", col = "dodgerblue3") +
theme(panel.background = element_rect(fill = "white"),
axis.line.x=element_line(),
axis.line.y=element_line()) +
ggtitle("Linear Model Fitted to Case Data")

deathPred <- round(predict(fit_1, data.frame(index = 355)),0)
casePred <- round(predict(fit_2, data.frame(index = 355)),0)

sprintf("Based on my model, 5 days after the lasts datapoint, the number of deaths will be %f and the number of cases will be %f",deathPred,casePred)

```
```{r}
#Final report: Submit a pdf file with your analysis results and conclusions.

print("Overall I was able to build a model that was relatively well fit. The R-Squared, F-Statistic, and p-value for the cases prediction seemed to support a relatively well fit linear #regression. On the Deaths, the previosly mentioned measurements were a little less convincing, but overall the line of best fit seemed to provide valuable insight to a positively #sloped regression.")

```


